### 2019-06-04 numerics_paper/numerical_chj.tex
Add outline for numerical paper.

\documentclass{article}

\title{Numerical simulation of the shallow water circular hydraulic jump}

\author{
    David I. Ketcheson \and
    Manuel Quezada de Luna
}
\begin{document}

\maketitle

\begin{abstract}
We study numerical discretrizations of the shallow water equations
for the simulation of the circular hydraulic jump formed
when a vertical fluid jet impacts a horizontal plate.  
We focus on three numerical methods: finite volume Godunov-type
methods using either a Roe or HLLE approximate Riemann solver, and
a continuous Galerkin finite element method based on flux-corrected transport (FCT).
The finite volume methods are employed ona uniform cartesian grid
while the finite element method uses an unstructured triangular grid.
The three methods give completely different results, even for
highly-refined meshes.
\end{abstract}

\section{Introduction}

\section{The shallow water circular hydraulic jump}

\subsection{Semi-analytical solution under rotational symmetry}

\section{Numerical methods}

\subsection{Finite volume methods}

\subsubsection{Roe solver}

\subsubsection{HLLE solver}

\subsection{FCT finite element method}

\section{Numerical results}

\section{Conclusions}

\end{document}