### 2020-06-11 numerics_paper/numerical_chj.tex
Introductory material on the CHJ.

\documentclass{article}

\title{Numerical simulation of the shallow water circular hydraulic jump}

\author{
    David I. Ketcheson \and
    Manuel Quezada de Luna
}
\begin{document}

\maketitle

\begin{abstract}
We study numerical discretrizations of the shallow water equations
for the simulation of the circular hydraulic jump formed
when a vertical fluid jet impacts a horizontal plate.  
We focus on three numerical methods: finite volume Godunov-type
methods using either a Roe or HLLE approximate Riemann solver, and
a continuous Galerkin finite element method based on flux-corrected transport (FCT).
The finite volume methods are employed ona uniform cartesian grid
while the finite element method uses an unstructured triangular grid.
The three methods give completely different results, even for
highly-refined meshes.
\end{abstract}

\section{Introduction}

\subsection{The circular hydraulic jump}

Perhaps the first reference to the observation of the circular hydraulic jump
comes from Lord Rayleigh \cite{rayleigh1914theory}, who wrote that it
``may usually be seen whenever a stream of water from a tap strikes a horizontal
surface".  This phenomenon that is familiar in the everyday kitchen sink, is in
fact highly nonlinear and unintuitive.  Near the jet, the flow is very shallow and
fast, while further away it is deeper and slower.  The transition between these regions
occurs in a very narrow region and takes the form of \emph{jump} or \emph{bore}
that is roughly circular if the surface is flat; we refer to it herein as a
circular hydraulic jump (CHJ).

Early experimental work the CHJ began some time later
\cite{kurihara1946hydraulic,tani1949water,watson1964radial}.
Watson \cite{watson1964radial} worked out the jump radius
implied by Rayleigh's approach and then worked out the vertical
velocity profile in the inner region, assuming a no-slip
boundary condition at the bottom.  He also studied the turbulent
flow case and performed experiments.
More detailed experiments revealed different qualitative classification
of jumps  \cite{ishigai1977heat,craik1981circular}.
Although later work incorporated more physical details (such as surface tension) into the models
\cite{bush2003influence}, Bohr et. al. showed that important properties of the jump
(particularly its radius) could be reasonably predicted using a simple shallow water
model \cite{bohr1993shallow}.

While the jump is roughly circular, it may deviate from this shape and
deforming rapidly in time.
Instability of the jump was observed from fairly early on \cite{craik1981circular}.
Under special circumstances with more viscous fluids, the jump instability may lead to
the formation of curious shapes such as polygons \cite{ellegaard1998creating}, but
for a low-viscosity fluid like water the behavior is generally chaotic.
The instability increases with the jet velocity and with the depth at the outside of the jump,
and can be stable for sufficiently small velocities and depths.

\subsection{Numerical shock instabilities}

\section{The shallow water circular hydraulic jump}

\subsection{Semi-analytical solution under rotational symmetry}

\section{Numerical methods}

\subsection{Finite volume methods}

\subsubsection{Roe solver}

\subsubsection{HLLE solver}

\subsubsection{HLLM solver}

\subsection{Continuous Galerkin method}

\subsection{Discontinuous Galerkin method}

\section{Numerical results}

\subsection{Stable jump}

\subsection{Unstable jump}

\section{Conclusions}

\bibliographystyle{apalike}
\bibliography{refs}

\end{document}